---
title: "FN Peru-only model reference-point diagnostics"
date: today
format:
  html:
    output-file: "FN-ReferencePoints.html"
    self-contained: true
    lightbox: true
    html-math-method: katex
    toc: true
    toc-title: "Contents"
    toc-depth: 2
    toc-location: left
    number-sections: true
    number-depth: 2
    theme: cosmo
    fig-width: 9
    fig-height: 5
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(jjmR)
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)

date_stamp <- format(Sys.time(), "Rendered %d %B %Y, %H:%M %Z")
stock_id <- 1

doc_dir <- dirname(knitr::current_input(dir = TRUE))
assessment_candidates <- c(
  Sys.getenv("JJM_ASSESSMENT_DIR", unset = ""),
  file.path(doc_dir, "..", "assessment"),
  file.path(doc_dir, "..", "..", "jjm", "assessment")
)
assessment_candidates <- assessment_candidates[nzchar(assessment_candidates)]
assessment_dir <- assessment_candidates[file.exists(file.path(assessment_candidates, "config"))][1]
if (is.na(assessment_dir)) {
  stop("Could not find the JJM assessment directory. Set JJM_ASSESSMENT_DIR to its path.")
}
assessment_dir <- normalizePath(assessment_dir, winslash = "/", mustWork = TRUE)
assessment_file <- function(...) file.path(assessment_dir, ...)
source(assessment_file("R", "diagnostic_composition_plots.R"))

model_info <- tibble(
  model = c("FN", "FN_msy_sel5"),
  label = c("FN base", "FN -msy_sel 5"),
  msy_sel_argument = c("default", "-msy_sel 5")
) |>
  mutate(label = factor(label, levels = label))

model_colors <- c("FN base" = "#2c7fb8", "FN -msy_sel 5" = "#d95f0e")

read_model_bundle <- function(model_name) {
  readJJM(
    model_name,
    path = assessment_file("config"),
    input = assessment_file("input"),
    output = assessment_file("results"),
    version = "2015MS"
  )
}

read_yield_profile <- function(model_name) {
  read.table(
    assessment_file("results", paste0(model_name, ".yld")),
    skip = 5,
    fill = TRUE,
    col.names = c("f", "spawning_biomass", "yield", "recruitment", "spr", "total_biomass")
  ) |>
    as_tibble() |>
    mutate(across(everything(), as.numeric)) |>
    filter(!is.na(f), !is.na(spawning_biomass), !is.na(yield), !is.na(spr))
}

read_std_terms <- function(model_name, terms, occurrence = 1) {
  std_file <- assessment_file("results", paste0(model_name, ".std"))
  out <- tibble(term = terms, value = NA_real_, std_dev = NA_real_)
  if (!file.exists(std_file)) {
    return(out)
  }

  std_output <- read.table(std_file, header = TRUE)
  matched <- vapply(terms, function(term) {
    matches <- which(std_output$name == term)
    if (length(matches) < occurrence) {
      NA_integer_
    } else {
      matches[[occurrence]]
    }
  }, integer(1))

  out$value <- as.numeric(std_output$value[matched])
  out$std_dev <- as.numeric(std_output$std.dev[matched])
  out
}

read_std_series <- function(model_name, term, years) {
  std_file <- assessment_file("results", paste0(model_name, ".std"))
  if (!file.exists(std_file)) {
    return(tibble(year = years, value = NA_real_, std_dev = NA_real_))
  }

  std_output <- read.table(std_file, header = TRUE)
  matches <- which(std_output$name == term)
  matches <- matches[seq_len(min(length(matches), length(years)))]

  tibble(
    year = years[seq_along(matches)],
    value = as.numeric(std_output$value[matches]),
    std_dev = as.numeric(std_output$std.dev[matches])
  )
}

read_r_report_value <- function(model_name, term, stock = 1) {
  r_report_file <- assessment_file("results", paste0(model_name, "_", stock, "_R.rep"))
  if (!file.exists(r_report_file)) {
    return(NA_real_)
  }

  r_report <- readLines(r_report_file, warn = FALSE)
  term_line <- which(r_report == paste0("$", term))[1]
  if (is.na(term_line) || term_line == length(r_report)) {
    return(NA_real_)
  }

  as.numeric(strsplit(trimws(r_report[term_line + 1]), "\\s+")[[1]][1])
}

read_par_summary <- function(model_name) {
  par_file <- assessment_file("results", paste0(model_name, ".par"))
  vals <- as.numeric(scan(par_file, what = "", n = 16, quiet = TRUE)[c(6, 11, 16)])
  tibble(n_parameters = as.integer(vals[1]), nll = vals[2], max_gradient = vals[3])
}

interpolate_spr <- function(profile, target_spr = 0.35) {
  profile <- arrange(profile, f)
  crossing <- which(
    (head(profile$spr, -1) - target_spr) *
      (tail(profile$spr, -1) - target_spr) <= 0
  )[1]

  if (is.na(crossing)) {
    return(tibble(
      profile_f_at_spr35 = NA_real_,
      sbf35 = NA_real_,
      yield_at_spr35 = NA_real_,
      recruitment_at_spr35 = NA_real_,
      total_biomass_at_spr35 = NA_real_
    ))
  }

  lower <- profile[crossing, ]
  upper <- profile[crossing + 1, ]
  weight <- (target_spr - lower$spr) / (upper$spr - lower$spr)

  tibble(
    profile_f_at_spr35 = lower$f + weight * (upper$f - lower$f),
    sbf35 = lower$spawning_biomass + weight * (upper$spawning_biomass - lower$spawning_biomass),
    yield_at_spr35 = lower$yield + weight * (upper$yield - lower$yield),
    recruitment_at_spr35 = lower$recruitment + weight * (upper$recruitment - lower$recruitment),
    total_biomass_at_spr35 = lower$total_biomass + weight * (upper$total_biomass - lower$total_biomass)
  )
}

interpolate_f <- function(profile, target_f) {
  if (!is.finite(target_f)) {
    return(tibble(
      spawning_biomass_at_f = NA_real_,
      yield_at_f = NA_real_,
      recruitment_at_f = NA_real_,
      spr_at_f = NA_real_,
      total_biomass_at_f = NA_real_
    ))
  }

  profile <- arrange(profile, f)
  tibble(
    spawning_biomass_at_f = approx(profile$f, profile$spawning_biomass, xout = target_f)$y,
    yield_at_f = approx(profile$f, profile$yield, xout = target_f)$y,
    recruitment_at_f = approx(profile$f, profile$recruitment, xout = target_f)$y,
    spr_at_f = approx(profile$f, profile$spr, xout = target_f)$y,
    total_biomass_at_f = approx(profile$f, profile$total_biomass, xout = target_f)$y
  )
}

as_estimate_df <- function(x, quantity = NULL) {
  out <- as_tibble(x)
  names(out) <- c("year", "estimate", "sd", "lower", "upper")
  out <- mutate(out, across(everything(), as.numeric))
  if (!is.null(quantity)) {
    out <- mutate(out, quantity = quantity, .before = 1)
  }
  out
}

theme_report <- function() {
  theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      panel.grid.minor = element_blank(),
      plot.title.position = "plot"
    )
}

model_bundles <- setNames(lapply(model_info$model, read_model_bundle), model_info$model)

summarize_model <- function(model_name, label, msy_sel_argument) {
  model <- model_bundles[[model_name]][[1]]
  stock_output <- model$output[[stock_id]]
  profile <- read_yield_profile(model_name)
  profile_positive <- profile |>
    filter(spawning_biomass >= 0, yield >= 0, recruitment >= 0)
  profile_msy_row <- profile_positive |>
    slice_max(yield, n = 1, with_ties = FALSE)
  std_terms <- read_std_terms(model_name, c("Fmsy", "Fcur_Fmsy", "Bmsy", "Bcur_Bmsy"))
  reported_fmsy <- std_terms$value[std_terms$term == "Fmsy"]
  fmsy <- if (is.finite(reported_fmsy)) reported_fmsy else profile_msy_row$f
  f35_est <- read_r_report_value(model_name, "F35_est", stock = stock_id)
  if (!is.finite(f35_est)) {
    f35_est <- as.numeric(stock_output$F35_est)
  }

  spr35 <- interpolate_spr(profile_positive, target_spr = 0.35)
  fmsy_profile_values <- interpolate_f(profile_positive, fmsy)

  tibble(
    model = model_name,
    label = label,
    msy_sel_argument = msy_sel_argument,
    stock = stock_id,
    F35_est = f35_est,
    F_at_SPR35_profile = spr35$profile_f_at_spr35,
    Fmsy = fmsy,
    Fmsy_se = std_terms$std_dev[std_terms$term == "Fmsy"],
    profile_Fmsy = profile_msy_row$f,
    MSY = profile_msy_row$yield,
    SPR_at_Fmsy = fmsy_profile_values$spr_at_f,
    SBF35 = spr35$sbf35,
    yield_at_SPR35 = spr35$yield_at_spr35,
    yield_at_Fmsy = fmsy_profile_values$yield_at_f,
    Bmsy = std_terms$value[std_terms$term == "Bmsy"],
    Bmsy_se = std_terms$std_dev[std_terms$term == "Bmsy"],
    Fcur_Fmsy = std_terms$value[std_terms$term == "Fcur_Fmsy"],
    Bcur_Bmsy = std_terms$value[std_terms$term == "Bcur_Bmsy"]
  )
}

reference_summary <- bind_rows(
  Map(summarize_model, model_info$model, model_info$label, model_info$msy_sel_argument)
) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

run_summary <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  model_name <- model_info$model[[i]]
  model_obj <- model_bundles[[model_name]][[1]]
  stock_output <- model_obj$output[[stock_id]]

  tibble(
    model = model_name,
    label = model_info$label[[i]],
    msy_sel_argument = model_info$msy_sel_argument[[i]],
    data_file = model_obj$control$dataFile,
    stocks = model_obj$control$nStocks,
    fisheries = model_obj$data$Fnum,
    fishery_names = paste(model_obj$data$Fnames, collapse = ", "),
    indices = model_obj$data$Inum,
    index_names = paste(model_obj$data$Inames, collapse = ", "),
    fishery_age_comp_years = sum(model_obj$data$FnumyearsA),
    fishery_length_comp_years = sum(model_obj$data$FnumyearsL),
    end_year = max(as.integer(stock_output$Yr))
  ) |>
    bind_cols(read_par_summary(model_name))
})) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

yield_profiles <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  read_yield_profile(model_info$model[[i]]) |>
    filter(spawning_biomass >= 0, yield >= 0, recruitment >= 0) |>
    mutate(
      model = model_info$model[[i]],
      label = model_info$label[[i]],
      .before = 1
    )
})) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

profile_plot_max_f <- max(reference_summary$F35_est, reference_summary$Fmsy, reference_summary$profile_Fmsy, na.rm = TRUE) * 1.8
f_profile <- yield_profiles |>
  filter(f <= profile_plot_max_f) |>
  select(model, label, f, `Spawning biomass` = spawning_biomass, Yield = yield, Recruitment = recruitment, SPR = spr) |>
  pivot_longer(c(`Spawning biomass`, Yield, Recruitment, SPR), names_to = "quantity", values_to = "value") |>
  mutate(quantity = factor(quantity, levels = c("Yield", "Spawning biomass", "Recruitment", "SPR")))

f_reference_lines <- reference_summary |>
  select(model, label, F35_est, Fmsy) |>
  pivot_longer(c(F35_est, Fmsy), names_to = "reference", values_to = "f") |>
  mutate(
    reference = recode(reference, F35_est = "F35%", Fmsy = "Fmsy"),
    reference = factor(reference, levels = c("F35%", "Fmsy"))
  )

fbar_series <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  model_name <- model_info$model[[i]]
  years <- as.integer(model_bundles[[model_name]][[1]]$output[[stock_id]]$Yr)
  read_std_series(model_name, "Fbar", years) |>
    mutate(
      model = model_name,
      label = model_info$label[[i]],
      lower = pmax(value - 1.96 * std_dev, 0),
      upper = value + 1.96 * std_dev,
      .before = 1
    )
})) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

fishery_f <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  model_name <- model_info$model[[i]]
  stock_output <- model_bundles[[model_name]][[1]]$output[[stock_id]]
  out <- as_tibble(stock_output$F_fsh_1, .name_repair = "minimal")
  names(out) <- c("year", "mean_f", "max_selectivity_scaled_f")
  out |>
    mutate(across(everything(), as.numeric)) |>
    pivot_longer(-year, names_to = "quantity", values_to = "value") |>
    mutate(
      model = model_name,
      label = model_info$label[[i]],
      quantity = recode(
        quantity,
        mean_f = "Mean F",
        max_selectivity_scaled_f = "Max-selectivity-scaled F"
      ),
      .before = 1
    )
})) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

selectivity_at_age <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  model_name <- model_info$model[[i]]
  label <- model_info$label[[i]]
  stock_output <- model_bundles[[model_name]][[1]]$output[[stock_id]]

  terminal_selectivity <- suppressWarnings(get_selectivities(model_bundles[[model_name]])) |>
    filter(fleet_type == "fsh", fleet_name == "FarNorth", stock == "Stock_1") |>
    mutate(year = as.integer(year)) |>
    group_by(year) |>
    arrange(age, .by_group = TRUE) |>
    mutate(actual_age = row_number()) |>
    ungroup() |>
    filter(year == max(year)) |>
    transmute(
      model = model_name,
      label = label,
      age = actual_age,
      selectivity,
      source = "Terminal FarNorth"
    )

  msy_selectivity <- tibble(
    model = model_name,
    label = label,
    age = seq_along(as.numeric(stock_output$sel_msy)),
    selectivity = as.numeric(stock_output$sel_msy),
    source = "sel_msy"
  )

  bind_rows(terminal_selectivity, msy_selectivity)
})) |>
  mutate(
    label = factor(label, levels = levels(model_info$label)),
    source = factor(source, levels = c("Terminal FarNorth", "sel_msy"))
  )

biomass_series <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  model_name <- model_info$model[[i]]
  stock_output <- model_bundles[[model_name]][[1]]$output[[stock_id]]
  df1 <- as_tibble(stock_output$df1)
  names(df1) <- c("quantity", "year", "estimate", "sd", "lower", "upper")
  df1 |>
    mutate(across(-quantity, as.numeric)) |>
    filter(quantity %in% c("SSB", "SSB_Nofishing")) |>
    mutate(
      model = model_name,
      label = model_info$label[[i]],
      biomass_type = recode(
        quantity,
        SSB = "With fishing",
        SSB_Nofishing = "Dynamic Bzero"
      ),
      .before = 1
    )
})) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

depletion_series <- bind_rows(lapply(seq_len(nrow(model_info)), function(i) {
  model_name <- model_info$model[[i]]
  stock_output <- model_bundles[[model_name]][[1]]$output[[stock_id]]
  as_estimate_df(stock_output$SSB_NoFishR, "SSB_NoFishR") |>
    mutate(
      model = model_name,
      label = model_info$label[[i]],
      .before = 1
  )
})) |>
  mutate(label = factor(label, levels = levels(model_info$label)))

base_model_name <- "FN"
base_model_label <- "FN base"
base_bundle <- model_bundles[[base_model_name]]
base_model <- base_bundle[[1]]

extract_fn_index_fits <- function(mod_bundle, stock = 1) {
  mod <- mod_bundle[[1]]
  index_names <- trimws(as.character(mod$data$Inames))
  out <- mod$output[[stock]]
  survey_names <- grep("^Obs_Survey_[0-9]+$", names(out), value = TRUE)

  bind_rows(lapply(survey_names, function(survey_name) {
    index_id <- as.integer(sub("^Obs_Survey_([0-9]+)$", "\\1", survey_name))
    x <- as.matrix(out[[survey_name]])
    if (is.null(x) || ncol(x) < 3) {
      return(tibble())
    }

    tibble(
      index_id = index_id,
      index = if (index_id <= length(index_names)) index_names[[index_id]] else paste0("Index_", index_id),
      year = as.integer(x[, 1]),
      observed = as.numeric(x[, 2]),
      predicted = as.numeric(x[, 3]),
      sd = if (ncol(x) >= 4) as.numeric(x[, 4]) else NA_real_,
      std_residual = if (ncol(x) >= 5) as.numeric(x[, 5]) else NA_real_,
      log_std_residual = if (ncol(x) >= 6) as.numeric(x[, 6]) else NA_real_
    )
  })) |>
    mutate(index = gsub("_", " ", index))
}

fn_index_fits <- extract_fn_index_fits(base_bundle, stock = stock_id)

fn_catch_biomass <- tibble(
  year = as.integer(base_model$output[[stock_id]]$Yr),
  observed = as.numeric(base_model$output[[stock_id]]$Obs_catch_1),
  predicted = as.numeric(base_model$output[[stock_id]]$Pred_catch_1)
) |>
  pivot_longer(c(observed, predicted), names_to = "series", values_to = "catch_biomass") |>
  mutate(
    series = recode(series, observed = "Observed", predicted = "Predicted"),
    series = factor(series, levels = c("Observed", "Predicted"))
  )

fn_length_fits <- diagnostic_comp_data(base_bundle, stock = stock_id, comp_type = "length") |>
  filter(source_type == "fsh") |>
  mutate(
    period = case_when(
      year < 2002 ~ "1972-2001",
      year <= 2015 ~ "2002-2015",
      TRUE ~ "2016-2025"
    ),
    period = factor(period, levels = c("1972-2001", "2002-2015", "2016-2025"))
  )

fn_length_observed <- fn_length_fits |>
  group_by(source_name, year, bin) |>
  summarise(
    observed = first(observed),
    sample_size = first(sample_size),
    period = first(period),
    .groups = "drop"
  )

fn_selectivity_time <- suppressWarnings(get_selectivities(base_bundle)) |>
  filter(fleet_type == "fsh", fleet_name == "FarNorth", stock == "Stock_1") |>
  mutate(year = as.integer(year)) |>
  group_by(year) |>
  arrange(age, .by_group = TRUE) |>
  mutate(actual_age = row_number()) |>
  ungroup()

plot_fn_index_fits <- function() {
  ggplot(fn_index_fits, aes(year, predicted)) +
    geom_linerange(
      aes(ymin = pmax(observed - 1.96 * sd, 0), ymax = observed + 1.96 * sd),
      color = "grey45",
      alpha = 0.45,
      linewidth = 0.45,
      na.rm = TRUE
    ) +
    geom_point(aes(y = observed), color = "black", size = 1.6, alpha = 0.85, na.rm = TRUE) +
    geom_line(color = model_colors[[base_model_label]], linewidth = 0.9, na.rm = TRUE) +
    facet_wrap(vars(index), scales = "free_y", ncol = 1) +
    scale_x_continuous(breaks = scales::breaks_width(3)) +
    labs(x = "Year", y = "Index value") +
    theme_report()
}

plot_fn_catch_biomass <- function() {
  ggplot(fn_catch_biomass, aes(year, catch_biomass, color = series)) +
    geom_line(linewidth = 0.85, na.rm = TRUE) +
    geom_point(
      data = filter(fn_catch_biomass, series == "Observed"),
      size = 1.4,
      alpha = 0.85,
      na.rm = TRUE
    ) +
    scale_color_manual(values = c("Observed" = "black", "Predicted" = model_colors[[base_model_label]])) +
    scale_x_continuous(breaks = scales::breaks_width(5)) +
    labs(x = "Year", y = "Catch biomass", color = NULL) +
    theme_report()
}

plot_fn_length_aggregate <- function() {
  aggregate_fit <- fn_length_fits |>
    filter(is.finite(observed), is.finite(fitted)) |>
    mutate(weight = if_else(is.finite(sample_size) & sample_size > 0, sample_size, 1)) |>
    group_by(period, bin) |>
    summarise(
      fitted = sum(fitted * weight, na.rm = TRUE) / sum(weight, na.rm = TRUE),
      observed = sum(observed * weight, na.rm = TRUE) / sum(weight, na.rm = TRUE),
      .groups = "drop"
    )

  ggplot(aggregate_fit, aes(bin, fitted)) +
    geom_col(
      aes(y = observed),
      fill = "#c6dcea",
      color = "#3c88ad",
      linewidth = 0.25,
      width = 0.85,
      alpha = 0.9
    ) +
    geom_line(color = model_colors[[base_model_label]], linewidth = 0.95) +
    geom_point(color = model_colors[[base_model_label]], size = 1.3) +
    facet_wrap(vars(period), ncol = 1) +
    scale_x_continuous(breaks = scales::breaks_width(5)) +
    labs(x = "Length", y = "Proportion") +
    theme_report()
}

plot_fn_length_years <- function(year_min = NULL, ncol = 4) {
  dat <- fn_length_fits
  obs <- fn_length_observed
  if (!is.null(year_min)) {
    dat <- filter(dat, year >= year_min)
    obs <- filter(obs, year >= year_min)
  }

  ggplot(dat, aes(bin, fitted)) +
    geom_point(
      data = obs,
      aes(bin, observed),
      inherit.aes = FALSE,
      color = "black",
      size = 0.7,
      alpha = 0.72,
      na.rm = TRUE
    ) +
    geom_line(color = model_colors[[base_model_label]], linewidth = 0.55, alpha = 0.95, na.rm = TRUE) +
    facet_wrap(vars(year), ncol = ncol) +
    scale_x_continuous(breaks = scales::breaks_width(10)) +
    scale_y_continuous(limits = c(0, NA)) +
    labs(x = "Length", y = "Proportion") +
    theme_report()
}

plot_fn_selectivity_time <- function() {
  year_breaks <- sort(unique(fn_selectivity_time$year))
  year_breaks <- year_breaks[year_breaks %% 5 == 0]

  ggplot(
    fn_selectivity_time,
    aes(
      x = actual_age,
      y = forcats::fct_rev(factor(year)),
      height = selectivity,
      group = year
    )
  ) +
    ggridges::geom_density_ridges(
      stat = "identity",
      fill = model_colors[[base_model_label]],
      color = model_colors[[base_model_label]],
      alpha = 0.45,
      scale = 3.5,
      linewidth = 0.3
    ) +
    scale_x_continuous(breaks = 1:12, limits = c(1, 12)) +
    scale_y_discrete(breaks = as.character(year_breaks)) +
    labs(x = "Age", y = "Year") +
    theme_report() +
    theme(legend.position = "none")
}
```

::: {.content-visible .no-para-number unless-format="pdf"}
::: {.no-para-number style="text-align: center; margin-top: 0.75rem; margin-bottom: 1rem;"}
`r date_stamp`
:::
:::

# Purpose

This document is a focused diagnostic for the `FN` Peru-only test model. Both runs use the same single-stock configuration with FarNorth catch, FarNorth length composition data from `0.15.dat`, and the two Peruvian CPUE indices (`Peru_Artis` and `Peru_Ind`). The comparison isolates the command-line MSY selectivity option: the base run uses the default `jjm` behavior and the alternative run uses `-msy_sel 5`.

```{r}
#| label: tbl-run-summary
#| tbl-cap: "FN run configurations read from the active assessment outputs."

kable(run_summary, digits = 5)
```

# Base Run Fits

The diagnostics in this section use only the base `FN` run. The `FN_msy_sel5` run is omitted here because it was included only to evaluate the effect of the MSY selectivity option on reference-point calculations.

## Catch Biomass

The catch biomass panel compares the observed FarNorth catch biomass from the active data file with the fitted catch biomass from the base `FN` run.

```{r}
#| label: fig-fn-base-catch-biomass
#| fig-cap: "Observed and predicted FarNorth catch biomass by year for the base FN run."
#| fig-width: 8
#| fig-height: 5

plot_fn_catch_biomass()
```

## Index Fits

Points and vertical intervals show the observed Peruvian CPUE values and nominal 95% intervals from the active data file. The line shows the fitted values from the base `FN` run.

```{r}
#| label: fig-fn-base-index-fits
#| fig-cap: "Base FN fit to the Peruvian artisanal and industrial CPUE indices."
#| fig-width: 8
#| fig-height: 6

plot_fn_index_fits()
```

## Catch-At-Length Fits

The FarNorth catch-at-length panels show observed length proportions as bars or black points and fitted proportions as the blue line from the base `FN` run.

::: {.panel-tabset}

### Aggregate By Period

```{r}
#| label: fig-fn-base-length-aggregate
#| fig-cap: "Sample-size-weighted aggregate FarNorth catch-at-length fits for the base FN run."
#| fig-width: 8
#| fig-height: 7

plot_fn_length_aggregate()
```

### Recent Years

```{r}
#| label: fig-fn-base-length-recent
#| fig-cap: "Recent annual FarNorth catch-at-length fits for the base FN run."
#| fig-width: 10
#| fig-height: 9

plot_fn_length_years(year_min = 2002, ncol = 4)
```

### All Years

```{r}
#| label: fig-fn-base-length-all-years
#| fig-cap: "All annual FarNorth catch-at-length fits for the base FN run."
#| fig-width: 10
#| fig-height: 17

plot_fn_length_years(ncol = 4)
```

:::

## Selectivity Time Series

The ridge plot shows the fitted FarNorth fishery selectivity-at-age trajectory from the base `FN` run. Ages are plotted as actual model ages 1-12.

```{r}
#| label: fig-fn-base-selectivity-ridges
#| fig-cap: "Base FN FarNorth fishery selectivity-at-age time series. Ages are shown from 1 to 12."
#| fig-width: 8
#| fig-height: 9

plot_fn_selectivity_time()
```

# Reference Points

The runs have the same estimated fit but different MSY selectivity settings. As a result, the reference-point outputs differ even where the model likelihood and fitted population trajectory are unchanged.

```{r}
#| label: tbl-fn-reference-points
#| tbl-cap: "Reference-point summary for the FN base and -msy_sel 5 runs."

reference_summary |>
  kable(digits = 5)
```

```{r}
#| label: fig-fn-f-profile
#| fig-cap: "FN yield-profile quantities. Vertical lines mark F35% and Fmsy for each run."
#| fig-width: 9
#| fig-height: 6.5

ggplot(f_profile, aes(f, value, color = label)) +
  geom_line(linewidth = 0.85) +
  geom_vline(
    data = f_reference_lines,
    aes(xintercept = f, color = label, linetype = reference),
    linewidth = 0.85
  ) +
  facet_wrap(vars(quantity), scales = "free_y", ncol = 2) +
  scale_color_manual(values = model_colors) +
  scale_linetype_manual(values = c("F35%" = "dashed", "Fmsy" = "dotdash")) +
  labs(x = "F", y = NULL, color = "Run", linetype = "Reference point") +
  theme_report()
```

# Fishing Mortality

Annual $F_{bar}$ is the mean total fishing mortality over ages 2-5, matching the convention used in Paper 02. Only the base `FN` run is shown in this section. The interval uses the reported standard error for `Fbar` from the `.std` output.

```{r}
#| label: fig-fn-fbar
#| fig-cap: "Base FN annual Fbar over ages 2-5. Ribbons show approximate 95% intervals from the reported standard errors."

ggplot(filter(fbar_series, model == base_model_name), aes(year, value)) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper),
    fill = model_colors[[base_model_label]],
    alpha = 0.12,
    linewidth = 0
  ) +
  geom_line(color = model_colors[[base_model_label]], linewidth = 0.85) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  labs(x = "Year", y = "Mean total F ages 2-5") +
  theme_report()
```

```{r}
#| label: fig-fn-fishery-f
#| fig-cap: "Base FN FarNorth fishery annual fishing mortality from the model output."

ggplot(filter(fishery_f, model == base_model_name), aes(year, value)) +
  geom_line(color = model_colors[[base_model_label]], linewidth = 0.85) +
  facet_wrap(vars(quantity), scales = "free_y", ncol = 1) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  labs(x = "Year", y = "Fishing mortality") +
  theme_report()
```

# Selectivity

The terminal FarNorth selectivity comes from the fitted model output. The `sel_msy` vector is the selectivity vector used in the MSY calculations and is where the `-msy_sel 5` option changes the reference-point calculation.

```{r}
#| label: fig-fn-selectivity
#| fig-cap: "FN selectivity-at-age for the terminal FarNorth fishery and the MSY selectivity vector. Ages are shown as actual model ages 1-12."

ggplot(selectivity_at_age, aes(age, selectivity, color = label, linetype = source, shape = source)) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:12, limits = c(1, 12)) +
  scale_color_manual(values = model_colors) +
  scale_linetype_manual(values = c("Terminal FarNorth" = "solid", "sel_msy" = "dashed")) +
  labs(x = "Age", y = "Selectivity", color = "Run", linetype = "Source", shape = "Source") +
  theme_report()
```

# Biomass And Depletion

The biomass panel shows fished spawning biomass and the no-fishing spawning biomass trajectory reported by the model. The depletion panel uses `SSB_NoFishR` directly from the output. These panels are included to confirm whether the MSY selectivity argument affected only the reference-point calculations or also changed the fitted model trajectory.

```{r}
#| label: fig-fn-biomass-dynamic-bzero
#| fig-cap: "FN spawning biomass with fishing and dynamic Bzero."

ggplot(biomass_series, aes(year, estimate, color = label, fill = label)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.10, linewidth = 0) +
  geom_line(linewidth = 0.85) +
  facet_wrap(vars(biomass_type), scales = "free_y", ncol = 1) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  labs(x = "Year", y = "Spawning biomass", color = "Run", fill = "Run") +
  theme_report()
```

```{r}
#| label: fig-fn-depletion
#| fig-cap: "FN SSB depletion relative to dynamic Bzero from the reported SSB_NoFishR output."

ggplot(depletion_series, aes(year, estimate, color = label, fill = label)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.10, linewidth = 0) +
  geom_line(linewidth = 0.85) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "gray35", linewidth = 0.65) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  coord_cartesian(ylim = c(0, NA)) +
  labs(x = "Year", y = expression(SSB / SSB[F == 0]), color = "Run", fill = "Run") +
  theme_report()
```
