---
title: "SCW17/Paper-07 Projection evaluations"
subtitle: "Forward projection scenarios and selectivity sensitivity from jmMSE"
date: today
format:
  html:
    output-file: "SCW17-Paper-07-Projection-evaluations.html"
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    lightbox: true
execute:
  echo: true
  warning: false
  message: false
params:
  jmMSE_path: "../../jmMSE"
---

```{r}
#| label: setup
library(ggplot2)
library(knitr)
library(patchwork)

jmMSE_path <- params$jmMSE_path
cache_file <- file.path(jmMSE_path, "report", "projections", "cache.rda")
if(!file.exists(cache_file)) {
  stop("Projection cache not found at ", cache_file,
    ". Run source('run_projections.R') in the jmMSE repository before ",
    "rendering this report.")
}

load(cache_file)

fmt_num <- function(x, digits=0) {
  format(round(x, digits), big.mark=",", trim=TRUE, scientific=FALSE)
}

fmt_value <- function(x) {
  if(is.character(x)) x else fmt_num(x, digits=3)
}

scenario_tbl <- do.call(rbind, lapply(names(SCENARIOS), function(nm) {
  data.frame(
    scenario = nm,
    control_quantity = SCENARIOS[[nm]]$quant,
    target = fmt_value(SCENARIOS[[nm]]$value),
    stringsAsFactors = FALSE
  )
}))

refpt_tbl <- data.frame(
  stock = names(sbmsy_ref),
  SBMSY_t = fmt_num(unname(sbmsy_ref), 0),
  FMSY = fmt_num(unname(fmsy_ref), 3),
  stringsAsFactors = FALSE
)

block_tbl <- do.call(rbind, lapply(names(SELEX_BLOCKS), function(nm) {
  yrs <- SELEX_BLOCKS[[nm]]
  data.frame(
    block = nm,
    years = paste(min(yrs), max(yrs), sep="-"),
    stringsAsFactors = FALSE
  )
}))
```

# Purpose

This report documents the forward-projection outputs generated by `run_projections.R` and assembled by `report_run_projections.R` in the jmMSE repository.
It is a Quarto version of the cached HTML report: it loads the saved projection cache, summarizes the run settings, and displays the diagnostic plots that were produced during the projection workflow.

No projection calculations are rerun in this document.
The report uses the cached objects in `r cache_file`, which are produced by `run_projections.R`.
This keeps the report rendering step separate from the computational projection step.

# Workflow Summary

The projection workflow does the following:

- loads the selected operating model (`REF_OM`) and subsets it to `N_ITER` iterations;
- extends the model from the last historical year to the final projection year;
- calculates reference points and current-state summaries;
- evaluates user-defined fixed fishing mortality and fixed catch scenarios;
- compares terminal stock status across scenarios;
- explores sensitivity to future selectivity assumptions using rolling terminal-year windows; and
- compares named historical selectivity blocks as alternative frozen future selectivity patterns.

`report_run_projections.R` then reads the cache and writes `report/projections/projections_report.html` in jmMSE.
This Quarto file reads the same cache and presents the same report content in a format that can be rendered, edited, and extended with standard Quarto tooling.

# Run Summary

```{r}
#| label: tbl-run-settings
#| tbl-cap: "Projection run settings stored in the cached projection report."
settings_tbl <- data.frame(
  setting = c(
    "Operating model",
    "Iterations",
    "Projection horizon",
    "Selectivity averaging windows",
    "Named selectivity blocks"
  ),
  value = c(
    REF_OM,
    fmt_num(N_ITER),
    paste(INIT_YEAR + 1, FINAL_YEAR, sep="-"),
    paste(SELEX_WINDOWS, collapse=", "),
    paste(names(SELEX_BLOCKS), collapse=", ")
  ),
  stringsAsFactors = FALSE
)

kable(settings_tbl)
```

```{r}
#| label: tbl-reference-points
#| tbl-cap: "Reference points used to interpret projected stock status."
kable(refpt_tbl, col.names=c("Stock", "SBMSY (t)", "FMSY"))
```

For the first stock in the operating model, the cached current-state values are:

```{r}
#| label: tbl-current-state
#| tbl-cap: "Cached current-state values for the first stock in the operating model."
current_tbl <- data.frame(
  quantity = c(
    paste0("SSB in ", INIT_YEAR, " (t)"),
    paste0("SSB / SBMSY in ", INIT_YEAR),
    paste0("Fbar in ", INIT_YEAR),
    paste0("Catch in ", INIT_YEAR, " (t)")
  ),
  value = c(
    fmt_num(ssb2025),
    fmt_num(ssb2025 / sbmsy_ref[[1]], 2),
    fmt_num(f2025, 3),
    fmt_num(catch2025)
  ),
  stringsAsFactors = FALSE
)

kable(current_tbl)
```

# Scenario Definitions

The fixed projection scenarios are defined in `run_projections.R`.
Each scenario either fixes fishing mortality (`fbar`) or total catch (`catch`) over the projection years.
Special targets such as `FMSY`, `F2025`, and `C2025` are resolved during the projection run before the cache is saved.

```{r}
#| label: tbl-scenarios
#| tbl-cap: "User-defined fixed F and fixed catch scenarios."
kable(scenario_tbl, col.names=c("Scenario", "Control quantity", "Target"))
```

# Selectivity History

The first diagnostic summarizes historical catch selectivity at age by fishery and stock.
This context is important because future projections freeze or average historical selectivity patterns, so changes in recent selectivity can affect projected spawning biomass and catch.

```{r}
#| label: fig-selectivity-history
#| fig-cap: "Historical selectivity-at-age time series used to interpret the projection selectivity assumptions."
#| fig-width: 14
#| fig-height: 8
selex_ts_plot
```

# Historical Exploitation Metrics

The exploitation diagnostic compares alternative fishing mortality or harvest-rate summaries through the historical period.
These metrics provide context for the fixed-F scenarios and for the current fishing mortality level used by the `F2025` scenario.

```{r}
#| label: fig-exploitation-metrics
#| fig-cap: "Historical exploitation metrics for the selected operating model."
#| fig-width: 14
#| fig-height: 9
mets_plot
```

# Fixed F And Catch Scenarios

The main projection comparison applies the fixed-F and fixed-catch controls over the projection horizon.
The plot shows median trajectories by scenario for the cached projection summaries, including spawning biomass, fishing mortality, and catch where available.

```{r}
#| label: fig-fixed-scenarios
#| fig-cap: "Forward projections under the user-defined fixed F and fixed catch scenarios."
#| fig-width: 14
#| fig-height: 10
futs_plot
```

The terminal-status plot shows the distribution of projected spawning biomass relative to SBMSY in the final projection year.
This gives a compact comparison of scenario risk at the end of the projection horizon.

```{r}
#| label: fig-terminal-status
#| fig-cap: !expr paste("Distribution of SSB / SBMSY in", FINAL_YEAR, "under each projection scenario.")
#| fig-width: 12
#| fig-height: 6
status_plot
```

# Selectivity Window Sensitivity

The selectivity-window sensitivity analysis projects at FMSY while replacing future selectivity with averages over alternative terminal-year windows.
For example, a three-year window uses the mean selectivity pattern from the final three historical years.

```{r}
#| label: tbl-selectivity-windows
#| tbl-cap: "Terminal-year selectivity averaging windows."
window_tbl <- data.frame(
  window_years = SELEX_WINDOWS,
  description = paste("Mean selectivity over the final", SELEX_WINDOWS, "historical years")
)

kable(window_tbl, col.names=c("Window (years)", "Projection assumption"))
```

```{r}
#| label: fig-selectivity-window-sensitivity
#| fig-cap: "Sensitivity of projections to the terminal-year selectivity averaging window."
#| fig-width: 14
#| fig-height: 8
selex_plot
```

# Selectivity Block Sensitivity

The named-block sensitivity analysis is similar to the rolling-window analysis, but the averaged selectivity pattern comes from specified historical year blocks.
This is useful when historical periods represent different fishery or stock-selectivity regimes.

```{r}
#| label: tbl-selectivity-blocks
#| tbl-cap: "Named historical selectivity blocks used for projection sensitivity."
kable(block_tbl, col.names=c("Block", "Years"))
```

```{r}
#| label: fig-selectivity-block-sensitivity
#| fig-cap: "Sensitivity of projections to named historical selectivity blocks."
#| fig-width: 14
#| fig-height: 8
block_plot
```

# Outputs

The cached projection workflow writes two main files:

- `jmMSE/report/projections/cache.rda`, containing the settings, summary values, and plot objects used by this report; and
- `jmMSE/report/projections/projections_report.html`, the original HTML report built by `report_run_projections.R`.

This Quarto document provides an editable report source for the same projection results.
