---
title: "Model developments"
date: today
format:
  html:
    output-file: "SCW17-Paper-01-ModelDevelopments.html"
    self-contained: true
    lightbox: true
    html-math-method: katex
    toc: true
    toc-title: "Contents"
    toc-depth: 3
    toc-location: left
    number-sections: true
    number-depth: 3
    theme: cosmo
    fig-width: 9
    fig-height: 5
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
---

{{< include _includes/sprfmo-banner.html >}}

```{=html}
<div class="paper-title">SCW17/Paper-01: Model developments</div>
```

```{r}
#| label: setup
#| include: false

library(jjmR)
library(tidyverse)
library(kableExtra)
library(knitr)

date_stamp <- format(Sys.Date(), "%d %B %Y")
doc_dir <- dirname(knitr::current_input(dir = TRUE))
assessment_candidates <- c(
  Sys.getenv("JJM_ASSESSMENT_DIR", unset = ""),
  file.path(doc_dir, "..", "assessment"),
  file.path(doc_dir, "..", "..", "jjm", "assessment")
)
assessment_candidates <- assessment_candidates[nzchar(assessment_candidates)]
assessment_dir <- assessment_candidates[file.exists(file.path(assessment_candidates, "config"))][1]
if (is.na(assessment_dir)) {
  stop("Could not find the JJM assessment directory. Set JJM_ASSESSMENT_DIR to its path.")
}
assessment_dir <- normalizePath(assessment_dir, winslash = "/", mustWork = TRUE)
assessment_file <- function(...) file.path(assessment_dir, ...)
knitr::opts_knit$set(root.dir = assessment_dir)
theme_set(theme_jjm(base_size = 14))

core_versions <- c("0.00", "0.14", "0.15", "0.16")
srr_versions <- c("0.16hs", "0.16ls", "0.16ll", "0.16hl")
versions <- c(core_versions, srr_versions)
hypotheses <- c("h1", "h2")
core_model_ids <- as.vector(outer(hypotheses, core_versions, paste, sep = "_"))
srr_model_ids <- as.vector(outer(hypotheses, srr_versions, paste, sep = "_"))
model_ids <- c(core_model_ids, srr_model_ids)

model_descriptions <- tribble(
  ~model, ~model_label, ~role, ~description,
  "0.00", "0.00", "SCW16 baseline", "SCW16 baseline configuration retained to anchor the model-development bridge.",
  "0.14", "0.14", "SCW16 final bridge", "SCW16 proposed final configuration used as the main comparison point for subsequent development.",
  "0.15", "0.15", "Model development", "Uses the 0.15 data treatment, including separate nominal Peruvian artisanal and industrial CPUE indices. The Peruvian CPUE q random walks are removed.",
  "0.16", "0.16", "Productivity-regime development", "Same data treatment as 0.15. For h2, the change is in productivity-regime specifications: stock 2 is represented with one productivity regime rather than separate pre/post regimes. The h1_0.15.ctl and h1_0.16.ctl files have the same biological and data specifications.",
  "0.16hs", "0.16 HS", "Productivity exercise", "High steepness with short stock-recruitment years.",
  "0.16ls", "0.16 LS", "Productivity exercise", "Low steepness with short stock-recruitment years.",
  "0.16ll", "0.16 LL", "Productivity exercise", "Low steepness with long stock-recruitment years.",
  "0.16hl", "0.16 HL", "Productivity exercise", "High steepness with long stock-recruitment years."
)

display_model <- function(x) {
  labels <- setNames(model_descriptions$model_label, model_descriptions$model)
  out <- unname(labels[x])
  ifelse(is.na(out), x, out)
}

model_labels <- setNames(
  rep(model_descriptions$model_label, each = length(hypotheses)),
  as.vector(outer(hypotheses, model_descriptions$model, paste, sep = "_"))
)

after_tag <- function(lines, tag) {
  hit <- which(trimws(lines) == tag)
  if (!length(hit) || hit[1] == length(lines)) {
    return(NA_character_)
  }
  trimws(lines[hit[1] + 1])
}

read_ctl_summary <- function(model_id) {
  ctl_path <- assessment_file("config", paste0(model_id, ".ctl"))
  if (!file.exists(ctl_path)) {
    return(tibble(
      model_id = model_id,
      control_file = basename(ctl_path),
      data_file = NA_character_,
      model_name = NA_character_,
      nstocks = NA_character_,
      nregby_stock = NA_character_,
      rec_matrix = NA_character_,
      nyrs_sr = NA_character_,
      n_q = NA_integer_,
      projection_years = NA_character_,
      control_exists = FALSE
    ))
  }

  lines <- readLines(ctl_path, warn = FALSE)
  q_vals <- suppressWarnings(scan(text = after_tag(lines, "#qMatrix"), quiet = TRUE))
  tibble(
    model_id = model_id,
    control_file = basename(ctl_path),
    data_file = after_tag(lines, "#dataFile"),
    model_name = after_tag(lines, "#modelName"),
    nstocks = after_tag(lines, "#nStocks"),
    nregby_stock = after_tag(lines, "#nregbyStock"),
    rec_matrix = after_tag(lines, "#RecMatrix"),
    nyrs_sr = after_tag(lines, "#Nyrs_sr"),
    n_q = length(q_vals),
    projection_years = after_tag(lines, "#n_proj_yrs"),
    control_exists = TRUE
  )
}

result_status <- function(model_id) {
  tibble(
    model_id = model_id,
    has_par = file.exists(assessment_file("results", paste0(model_id, ".par"))),
    has_rep = file.exists(assessment_file("results", paste0(model_id, ".rep"))),
    has_std = file.exists(assessment_file("results", paste0(model_id, ".std"))),
    has_yld = file.exists(assessment_file("results", paste0(model_id, ".yld"))),
    has_r_report = file.exists(assessment_file("results", paste0(model_id, "_1_R.rep"))),
    completed_output = has_par & has_rep
  )
}

control_inventory <- map_dfr(model_ids, read_ctl_summary) |>
  left_join(map_dfr(model_ids, result_status), by = "model_id") |>
  mutate(
    hypothesis = str_remove(model_id, "_.*$"),
    model = str_remove(model_id, "^h[12]_"),
    model_label = display_model(model),
    .before = 1
  )

completed_versions <- versions[
  map_lgl(versions, ~ all(control_inventory$completed_output[control_inventory$model == .x]))
]
core_completed_versions <- intersect(core_versions, completed_versions)
srr_completed_versions <- intersect(srr_versions, completed_versions)
bridge_versions <- intersect(c("0.00", "0.14"), completed_versions)

read_model_output <- function(model_id) {
  tryCatch(
    readJJM(model_id, path = "config", input = "input"),
    error = function(e) NULL
  )
}

model_cache <- new.env(parent = emptyenv())

read_model <- function(model_id) {
  if (!exists(model_id, envir = model_cache, inherits = FALSE)) {
    assign(model_id, read_model_output(model_id), envir = model_cache)
  }
  mod <- get(model_id, envir = model_cache, inherits = FALSE)
  if (is.null(mod) || is.null(mod[[model_id]])) {
    return(NULL)
  }
  mod[[model_id]]
}

model_readable <- function(model_id) {
  !is.null(read_model(model_id))
}

par_summary <- function(model_id) {
  par_file <- assessment_file("results", paste0(model_id, ".par"))
  if (!file.exists(par_file)) {
    return(tibble(
      model_id = model_id,
      hypothesis = str_remove(model_id, "_.*$"),
      model = str_remove(model_id, "^h[12]_"),
      nopar = NA_integer_,
      nll = NA_real_,
      max_gradient = NA_real_
    ))
  }
  vals <- suppressWarnings(as.numeric(scan(par_file, what = "", n = 16, quiet = TRUE)[c(6, 11, 16)]))
  tibble(
    model_id = model_id,
    hypothesis = str_remove(model_id, "_.*$"),
    model = str_remove(model_id, "^h[12]_"),
    nopar = as.integer(vals[1]),
    nll = vals[2],
    max_gradient = vals[3]
  )
}

value_at_year <- function(x, year) {
  if (is.matrix(x) || is.data.frame(x)) {
    x <- as.matrix(x)
    val <- as.numeric(x[x[, 1] == year, 2])
    if (length(val) == 0) return(NA_real_)
    return(val)
  }
  NA_real_
}

terminal_model_metrics <- function(model_id) {
  mod <- read_model(model_id)
  if (is.null(mod)) {
    return(tibble())
  }
  endyr <- mod$data$years[1, 2]

  map_dfr(names(mod$output), function(stock) {
    out <- mod$output[[stock]]
    terminal_row <- endyr - mod$data$years[1, 1] + 1
    tibble(
      model_id = model_id,
      hypothesis = str_remove(model_id, "_.*$"),
      model = str_remove(model_id, "^h[12]_"),
      model_label = display_model(str_remove(model_id, "^h[12]_")),
      stock = stock,
      year = endyr,
      ssb = value_at_year(out$SSB, endyr),
      total_biomass = value_at_year(out$TotBiom, endyr),
      recruitment = value_at_year(out$R, endyr),
      max_total_f_at_age = max(as.numeric(out$TotF[terminal_row, ]), na.rm = TRUE)
    )
  })
}

comparison_delta <- function(metrics, base_model = "0.14") {
  if (!nrow(metrics) || !base_model %in% metrics$model) {
    return(tibble())
  }
  metrics |>
    select(model, model_label, hypothesis, stock, year, ssb, total_biomass, recruitment, max_total_f_at_age) |>
    left_join(
      metrics |>
        filter(model == base_model) |>
        select(
          hypothesis,
          stock,
          ssb_base = ssb,
          total_biomass_base = total_biomass,
          recruitment_base = recruitment,
          max_total_f_at_age_base = max_total_f_at_age
        ),
      by = c("hypothesis", "stock")
    ) |>
    mutate(
      ssb_delta_pct = 100 * (ssb / ssb_base - 1),
      total_biomass_delta_pct = 100 * (total_biomass / total_biomass_base - 1),
      recruitment_delta_pct = 100 * (recruitment / recruitment_base - 1),
      max_total_f_at_age_delta_pct = 100 * (max_total_f_at_age / max_total_f_at_age_base - 1)
    )
}

relabel_models <- function(mods, labels) {
  current_names <- names(mods)
  new_names <- unname(labels[current_names])
  names(mods) <- ifelse(is.na(new_names), current_names, new_names)
  mods
}

safe_compare_models <- function(model_names) {
  tryCatch(
    relabel_models(compareModels(model_names), model_labels),
    error = function(e) NULL
  )
}

plot_compare_or_note <- function(mods, what, main, note) {
  if (!is.null(mods)) {
    plot(mods, combine = TRUE, what = what, stack = FALSE, main = main)
    return(invisible(NULL))
  }

  plot.new()
  wrapped <- strwrap(note, width = 72)
  text(
    0.5,
    0.55,
    paste(wrapped, collapse = "\n"),
    cex = 0.9
  )
  invisible(NULL)
}

available_model_ids <- control_inventory |>
  filter(completed_output) |>
  pull(model_id)

readable_model_ids <- available_model_ids[map_lgl(available_model_ids, model_readable)]
readable_versions <- versions[
  map_lgl(versions, ~ all(paste0(hypotheses, "_", .x) %in% readable_model_ids))
]
core_readable_versions <- intersect(core_versions, readable_versions)
srr_readable_versions <- intersect(srr_versions, readable_versions)
bridge_plot_versions <- intersect(c("0.00", "0.14"), readable_versions)
plot_read_note <- paste(
  "Comparison plot not available in this render."
)

fit_summary <- map_dfr(model_ids, par_summary) |>
  mutate(model_label = display_model(model))

core_available_model_ids <- available_model_ids[str_remove(available_model_ids, "^h[12]_") %in% core_versions]
terminal_metrics <- map_dfr(core_available_model_ids, terminal_model_metrics)
terminal_deltas <- comparison_delta(terminal_metrics, base_model = "0.14")
```

This working paper mirrors the SCW16 assessment-summary structure but narrows the model set to the configurations relevant to the SCW17 model-development bridge. The two completed reference points are the SCW16 baseline run (`0.00`) and the SCW16 proposed final run (`0.14`). Models `0.15` and `0.16` are included as control-file developments for the MSE transition.

The final model-development discussions from SCW16 were presented in [Jack Mackerel Assessment SCW16](https://sprfmo.github.io/JM_SCW_prep/SCW16.html). This document serves as a bridge between that benchmark assessment discussion and the SCW17 MSE workshop, focusing only on the benchmark anchor runs and the subsequent model-development controls needed for the MSE transition.

# Model Set

```{r}
#| label: tbl-model-set
#| results: asis
#| tbl-cap: "Model configurations included in this paper."

model_descriptions |>
  filter(model %in% core_versions) |>
  select(model_label, role, description) |>
  knitr::kable(
    col.names = c("Model", "Role", "Description"),
    align = "lll"
  ) |>
  kableExtra::kable_styling(full_width = FALSE)
```

The source controls are read from `r assessment_file("config")`. The rendered document can be run from the SCW17 repository or from the JJM assessment checkout; set `JJM_ASSESSMENT_DIR` if the assessment directory is not found automatically.

```{r}
#| label: tbl-control-inventory
#| results: asis
#| tbl-cap: "Control-file and output availability for the SCW17 model-development set."

control_inventory |>
  filter(model %in% core_versions) |>
  transmute(
    Hypothesis = hypothesis,
    Model = model_label,
    `Control file` = control_file,
    `Data file` = data_file,
    `Model name` = model_name,
    `Stocks` = nstocks,
    `Productivity regimes by stock` = nregby_stock,
    `Recruitment matrix` = rec_matrix,
    `SR years` = nyrs_sr,
    `Number of q terms` = n_q,
    `Has deterministic output` = completed_output
  ) |>
  knitr::kable(align = c(rep("l", 9), "r", "l")) |>
  kableExtra::kable_styling(full_width = FALSE, font_size = 12)
```

# Model Developments Since SCW16

The `0.00` configuration is retained as the SCW16 baseline bridge run. The `0.14` configuration is retained as the SCW16 proposed final model. Model `0.15` starts from the `0.14` line of development and changes the Peruvian CPUE treatment to use separate nominal artisanal and industrial CPUE indices. The Peruvian CPUE catchability random walks used in `0.14` are not estimated for those two nominal indices in `0.15`.

Model `0.16` uses the same data treatment as `0.15` but changes the productivity-regime specification. Under the two-stock hypothesis, `h2_0.15.ctl` retains the separate stock-2 productivity regimes and the associated regime shift used in the SCW16 model line, whereas `h2_0.16.ctl` combines stock 2 into one productivity regime. Under the single-stock hypothesis, `h1_0.15.ctl` and `h1_0.16.ctl` have the same biological and data specifications; the model-development distinction is therefore only meaningful for the two-stock productivity-regime specification.

# Fit Summaries

```{r}
#| label: tbl-fit-summary
#| results: asis
#| tbl-cap: "Negative log-likelihood and maximum gradient for available deterministic fits. Blank rows indicate controls that are present but not yet run."

fit_summary |>
  filter(model %in% core_versions) |>
  mutate(
    nll = round(nll, 3),
    max_gradient = signif(max_gradient, 4)
  ) |>
  select(hypothesis, model_label, model_id, nopar, nll, max_gradient) |>
  knitr::kable(
    col.names = c("Hypothesis", "Model", "Run", "N parameters", "NLL", "Max gradient"),
    align = "lllrrr"
  ) |>
  kableExtra::kable_styling(full_width = FALSE)
```

# Terminal-Year Metrics

```{r}
#| label: tbl-terminal-metrics
#| results: asis
#| tbl-cap: "Terminal-year metrics for available deterministic outputs."

if (nrow(terminal_metrics)) {
  terminal_metrics |>
    mutate(
      ssb = round(ssb),
      total_biomass = round(total_biomass),
      recruitment = round(recruitment),
      max_total_f_at_age = round(max_total_f_at_age, 3)
    ) |>
    select(hypothesis, model_label, stock, year, ssb, total_biomass, recruitment, max_total_f_at_age) |>
    knitr::kable(
      col.names = c("Hypothesis", "Model", "Stock", "Year", "SSB", "Total biomass", "Recruitment", "Max total F at age"),
      align = "llllrrrr"
    ) |>
    kableExtra::kable_styling(full_width = FALSE)
} else {
  knitr::kable(tibble(note = "No deterministic outputs were available."))
}
```

```{r}
#| label: tbl-terminal-delta
#| results: asis
#| tbl-cap: "Terminal-year percent difference relative to 0.14 for available deterministic outputs."

if (nrow(terminal_deltas)) {
  terminal_deltas |>
    mutate(across(ends_with("_delta_pct"), ~round(.x, 1))) |>
    select(hypothesis, model_label, stock, year, ends_with("_delta_pct")) |>
    knitr::kable(
      col.names = c("Hypothesis", "Model", "Stock", "Year", "SSB (%)", "Total biomass (%)", "Recruitment (%)", "Max total F at age (%)"),
      align = "llllrrrr"
    ) |>
    kableExtra::kable_styling(full_width = FALSE)
} else {
  knitr::kable(tibble(note = "Percent differences will populate after at least one comparison model and 0.14 have deterministic outputs."))
}
```

# 0.00 to 0.14 Bridge

The plots below reproduce the core SCW16 bridge contrast between the `0.00` baseline and the `0.14` proposed final model. They are included here so that subsequent `0.15` and `0.16` outputs can be interpreted relative to the accepted SCW16 bridge.

```{r}
#| label: bridge-setup

h1_bridge <- NULL
h2_bridge <- NULL

if (length(bridge_plot_versions) >= 2) {
  h1_bridge <- safe_compare_models(paste0("h1_", bridge_plot_versions))
  h2_bridge <- safe_compare_models(paste0("h2_", bridge_plot_versions))
}
```

## Single-Stock Hypothesis

::: {.panel-tabset}

### Biomass

```{r}
#| label: fig-bridge-h1-biomass
#| fig-cap: "Biomass estimates for the single-stock 0.00 to 0.14 bridge."

plot_compare_or_note(h1_bridge, "biomass", "Biomass", plot_read_note)
```

### Recruitment

```{r}
#| label: fig-bridge-h1-recruitment
#| fig-cap: "Recruitment estimates for the single-stock 0.00 to 0.14 bridge."

plot_compare_or_note(h1_bridge, "recruitment", "Recruitment", plot_read_note)
```

### Fishing Mortality

```{r}
#| label: fig-bridge-h1-fmort
#| fig-cap: "Fishing mortality estimates for the single-stock 0.00 to 0.14 bridge."

plot_compare_or_note(h1_bridge, "ftot", "Total Fishing Mortality", plot_read_note)
```

:::

## Two-Stock Hypothesis

::: {.panel-tabset}

### Biomass

```{r}
#| label: fig-bridge-h2-biomass
#| fig-cap: "Biomass estimates for the two-stock 0.00 to 0.14 bridge."
#| fig-height: 7

plot_compare_or_note(h2_bridge, "biomass", "Biomass", plot_read_note)
```

### Recruitment

```{r}
#| label: fig-bridge-h2-recruitment
#| fig-cap: "Recruitment estimates for the two-stock 0.00 to 0.14 bridge."

plot_compare_or_note(h2_bridge, "recruitment", "Recruitment", plot_read_note)
```

### Fishing Mortality

```{r}
#| label: fig-bridge-h2-fmort
#| fig-cap: "Fishing mortality estimates for the two-stock 0.00 to 0.14 bridge."

plot_compare_or_note(h2_bridge, "ftot", "Total Fishing Mortality", plot_read_note)
```

:::

# Development Comparisons

This section compares the completed deterministic outputs for the SCW16 anchor models and the subsequent `0.15` and `0.16` development runs.

```{r}
#| label: development-setup

h1_completed <- NULL
h2_completed <- NULL

if (length(core_readable_versions) >= 2) {
  h1_completed <- safe_compare_models(paste0("h1_", core_readable_versions))
  h2_completed <- safe_compare_models(paste0("h2_", core_readable_versions))
}
```

## Single-Stock Hypothesis

The single-stock `0.15` and `0.16` controls have the same biological and data specifications, so any single-stock difference between those labels would indicate a run-management or file-version difference rather than a model-structure difference.

```{r}
#| label: fig-development-h1-biomass
#| fig-cap: "Single-stock biomass for completed model-development outputs."

plot_compare_or_note(h1_completed, "biomass", "Biomass", plot_read_note)
```

```{r}
#| label: fig-development-h1-fmort
#| fig-cap: "Single-stock fishing mortality for completed model-development outputs."

plot_compare_or_note(h1_completed, "ftot", "Total Fishing Mortality", plot_read_note)
```

## Two-Stock Hypothesis

The two-stock development contrast is expected to be between `0.15` and `0.16`. Model `0.15` retains the separate stock-2 productivity regimes, while `0.16` combines the stock-2 productivity years into one regime. This contrast is intended to isolate the effect of productivity-regime specification after the nominal Peruvian CPUE data treatment is introduced.

```{r}
#| label: fig-development-h2-biomass
#| fig-cap: "Two-stock biomass for completed model-development outputs."
#| fig-height: 7

plot_compare_or_note(h2_completed, "biomass", "Biomass", plot_read_note)
```

```{r}
#| label: fig-development-h2-fmort
#| fig-cap: "Two-stock fishing mortality for completed model-development outputs."

plot_compare_or_note(h2_completed, "ftot", "Total Fishing Mortality", plot_read_note)
```

# Notes

## Fishing-mortality cap

The current JJM template includes a smooth penalty on high average fishing mortality. For each stock and model year, total fishing mortality is first summed across fisheries for that stock. The intermediate value `Fbar_tmp` is the mean total F over ages 2-5. With `Fcap = 3`, `eps = 0.05`, and `lambda_F = 10`, the penalty contribution is:

$$
\mathrm{excess} =
{1 \over 2}\left[(Fbar\_tmp - Fcap) +
\sqrt{(Fbar\_tmp - Fcap)^2 + eps^2}\right]
$$

$$
\mathrm{fpen} \mathrel{+}= \lambda_F \times \mathrm{excess}^2
$$

This behaves like a differentiable approximation to `10 * max(0, Fbar_tmp - 3)^2`. It is nearly zero below the cap, mild near `Fbar_tmp = 3`, and increasingly strong when a stock-year Fbar exceeds the cap. The form avoids a hard discontinuity in the objective function and is intended to discourage the unrealistically high terminal fishing-mortality spikes seen in some model-development trials without imposing a non-differentiable bound.

## Output status

At render time, completed deterministic outputs were available for the core bridge models `r paste(core_completed_versions, collapse = ", ")`. The `0.16` productivity-exercise runs are summarized separately below.

# Productivity Exercise To Do Later

The `0.16` `hs`, `ls`, `ll`, and `hl` runs are retained as a separate productivity exercise rather than part of the main model-development bridge. These controls follow the suffix convention used for the `0.14` stock-recruitment scenarios: the first letter identifies steepness (`h` = 0.85, `l` = 0.65) and the second letter identifies the stock-recruitment year set (`s` = 2001-2015, `l` = 1970-2022). For the two-stock controls, the stock-2 stock-recruitment year set is retained from `h2_0.16`, while steepness is changed consistently across recruitment regimes.

```{r}
#| label: tbl-productivity-controls
#| results: asis
#| tbl-cap: "Control-file and output availability for the later productivity exercise."

control_inventory |>
  filter(model %in% srr_versions) |>
  transmute(
    Hypothesis = hypothesis,
    Model = model_label,
    `Control file` = control_file,
    `Data file` = data_file,
    `Model name` = model_name,
    `Stocks` = nstocks,
    `Productivity regimes by stock` = nregby_stock,
    `Recruitment matrix` = rec_matrix,
    `SR years` = nyrs_sr,
    `Number of q terms` = n_q,
    `Has deterministic output` = completed_output
  ) |>
  knitr::kable(align = c(rep("l", 9), "r", "l")) |>
  kableExtra::kable_styling(full_width = FALSE, font_size = 12)
```

```{r}
#| label: tbl-productivity-fits
#| results: asis
#| tbl-cap: "Fit summaries for the later productivity exercise."

fit_summary |>
  filter(model %in% srr_versions) |>
  mutate(
    nll = round(nll, 3),
    max_gradient = signif(max_gradient, 4)
  ) |>
  select(hypothesis, model_label, model_id, nopar, nll, max_gradient) |>
  knitr::kable(
    col.names = c("Hypothesis", "Model", "Run", "N parameters", "NLL", "Max gradient"),
    align = "lllrrr"
  ) |>
  kableExtra::kable_styling(full_width = FALSE)
```

```{r}
#| label: srr-setup

h1_srr <- NULL
h2_srr <- NULL
srr_plot_versions <- c("0.16", srr_readable_versions)

if (length(srr_readable_versions) >= 1 && all(paste0("h1_", srr_plot_versions) %in% readable_model_ids)) {
  h1_srr <- safe_compare_models(paste0("h1_", srr_plot_versions))
}

if (length(srr_readable_versions) >= 1 && all(paste0("h2_", srr_plot_versions) %in% readable_model_ids)) {
  h2_srr <- safe_compare_models(paste0("h2_", srr_plot_versions))
}
```

## Single-Stock Hypothesis

```{r}
#| label: fig-srr-h1-biomass
#| fig-cap: "Single-stock biomass for the 0.16 productivity exercise."

plot_compare_or_note(h1_srr, "biomass", "Biomass", plot_read_note)
```

```{r}
#| label: fig-srr-h1-fmort
#| fig-cap: "Single-stock fishing mortality for the 0.16 productivity exercise."

plot_compare_or_note(h1_srr, "ftot", "Total Fishing Mortality", plot_read_note)
```

## Two-Stock Hypothesis

```{r}
#| label: fig-srr-h2-biomass
#| fig-cap: "Two-stock biomass for the 0.16 productivity exercise."
#| fig-height: 7

plot_compare_or_note(h2_srr, "biomass", "Biomass", plot_read_note)
```

```{r}
#| label: fig-srr-h2-fmort
#| fig-cap: "Two-stock fishing mortality for the 0.16 productivity exercise."

plot_compare_or_note(h2_srr, "ftot", "Total Fishing Mortality", plot_read_note)
```
